Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-12 bbbq_article.tex
Removed notes to self

\richel{Weak, add refs}
All pathogens can be detected by the foreign proteins on 
their (bacterial or fungal) cell walls or (viral) envelope.
For bacteria and fungi, this is the main mode of their detection,
as these do not infect a cell with their (foreign) DNA.
T and B cells express MHC-II to detect foreign peptide fragments.
An immune response is started when MHC-II detects a pathogen.

@@

It is helpful to be able to predict which peptides are immunogenic,
in, among others, vaccine development \richel{Reference here}. 
Already for a decade, synthesized peptide fragments are used
to experimentally determine which peptides
are immunogenic \richel{Reference here}.
This approach, however, is tedious and costly.
Due to this, software was written that allows for silico 
predictions, which are reliable in 
practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.

@@

When using DNA sequences, one can use a skewed rate
in non- versus synonymous mutation
to detect the signature of selection (\cite{murrell2015gene}).
Using AA sequences, however, has its advantages,
as its is closer to the actual phenotype
selection acts upon: DNA may never be transcribed to RNA,
or its RNA may never be translated (\cite{diz2012proteomics})
\richel{improve upon this point}.
When using a proteome in phylogenetic research,
we know that the majority of proteins are selected to just
maintain their function most of the time, where
is the time spans there is selection, only a few AAs
can actually increase the 'fitness' of the
protein (\cite{anisimova2009investigating}).
There, when generalizing the dynamics of mostly purifying selection (to maintain
a protein's function) and a short duration of positive selection,
those genes that are selected cannot be detected (\cite{yang2000statistical}).

@@

% The KS test to determine if MHC-I has the same percentage 
% of epitopes overlapping with viral TMHs, 
% compared to human TMHs \richel{$\mathcal{H}_{1,V}$},
% resulted in a p-value of \richel{unknown},
% which makes us [reject/accept] the hypothesis that these percentages
% are sampled from the same distribution.

% The KS test to determine if MHC-I has the same percentage 
% of epitopes overlapping with bacterial TMHs, 
% compared to human TMHs \richel{$\mathcal{H}_{1,B}$},
% resulted in a p-value of \richel{unknown},
% which makes us [reject/accept] the hypothesis that these percentages
% are sampled from the same distribution.

\begin{figure}[!htbp]
  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
  \caption{
    Percentage of MHC-I epitopes overlapping with TMHs
    for a human, viral and bacterial proteome.
    The horizontal lines indicate the percentage as expected by chance.
    \richel{Refer to a table with the numbers as expected by chance)}
    See table \ref{tab:tmh_binders_mhc1} for the TMH and epitope counts.
    \richel{Epitope prediction by EpitopePrediction.}.
    Note that for smaller proteomes a percentage of zero is likelier.
    \richel{use of top 2\% binders}
  }
  \label{fig:1}
\end{figure}

@@

\begin{figure}[!htbp]
  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
  \caption{
    Percentage of MHC-II epitopes overlapping with TMHs
    for a human, viral and bacterial proteome.
    The horizontal lines indicate the percentage as expected by chance.
    \richel{Refer to a table with the numbers as expected by chance)}
    \richel{Epitope prediction by NetMHCIIpan.}
    See table \ref{tab:tmh_binders_mhc2} for the TMH and epitope counts.
    Note that for smaller proteomes a percentage of zero is likelier.
    \richel{use of top 2\% binders}
  }
  \label{fig:2}
\end{figure}

@@

We assume that epitopes 
derived from TMHs are actually presented by MHC-I and MHC-II.
Our analysis (at subsection \ref{subsec:elution_studies}) 
finds that 1.4\% (for MHC-I) and 4.0% (for MHC-II) 
of the epitopes presented 
overlap with a TMH.
It is unsure if this a true finding or a false positive.
One source of noise would stem from imperfect predictions
by the topology prediction software used. 
We have counteracted this by using two different topology prediction
software and confirm these have similar results.
Two biological sources of noise
are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
and peptide fusion \cite{delong2016pathogenic},
two processes that both result in peptide fragments
that did not originate from genome directly.
Note however, that both elution studies have a
methodological bias caused by mass spectronomy,
as hydrophobic fragments are detected less 
by mass spectronomy \richel{Look up wording and ref in Bianchi et al 2017}.

@@

In this experiment we predicted epitopes that overlap with 
TMHs from a human, bacterial and viral proteome,
would these proteins be expressed in a human host.
Bacteria, however have different cell membranes and cell walls, 
hence different structural requirements for a TMH.
Both topology prediction tools were trained to recognize
human TMHs (\richel{Check}), thus we cannot be sure that
the transmembrane regions predicted in bacterial proteins
are actually part of an integral membrane protein.
For the purpose of this study, we assume the 
error in topology predictions to be unbiased way towards topology.
In other words: that a bacterial TMH is incorrectly
predicted to be absent just as often as it is incorrectly
predicted to be present elsewhere.

@@

\paragraph{Which reference proteome?}

See \url{https://github.com/richelbilderbeek/bbbq_article/issues/155}.

>>>>

All pathogens can be detected by the foreign proteins on 
their (bacterial or fungal) cell walls or (viral) envelope.
For bacteria and fungi, this is the main mode of their detection,
as these do not infect a cell with their (foreign) DNA.
T and B cells express MHC-II to detect foreign peptide fragments.
An immune response is started when MHC-II detects a pathogen.

@@

It is helpful to be able to predict which peptides are immunogenic,
in, among others, vaccine development. 
Already for a decade, synthesized peptide fragments are used
to experimentally determine which peptides
are immunogenic.
This approach, however, is tedious and costly.
Due to this, software was written that allows for silico 
predictions, which are reliable in 
practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.

@@

When using DNA sequences, one can use a skewed rate
in non- versus synonymous mutation
to detect the signature of selection (\cite{murrell2015gene}).
Using AA sequences, however, has its advantages,
as its is closer to the actual phenotype
selection acts upon: DNA may never be transcribed to RNA,
or its RNA may never be translated (\cite{diz2012proteomics}).
When using a proteome in phylogenetic research,
we know that the majority of proteins are selected to just
maintain their function most of the time, where
is the time spans there is selection, only a few AAs
can actually increase the 'fitness' of the
protein (\cite{anisimova2009investigating}).
There, when generalizing the dynamics of mostly purifying selection (to maintain
a protein's function) and a short duration of positive selection,
those genes that are selected cannot be detected (\cite{yang2000statistical}).

@@

\begin{figure}[!htbp]
  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
  \caption{
    Percentage of MHC-I epitopes overlapping with TMHs
    for a human, viral and bacterial proteome.
    The horizontal lines indicate the percentage as expected by chance.
    See table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
  }
  \label{fig:1}
\end{figure}

@@

\begin{figure}[!htbp]
  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
  \caption{
    Percentage of MHC-II epitopes overlapping with TMHs
    for a human, viral and bacterial proteome.
    The horizontal lines indicate the percentage as expected by chance.
    See table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
    Note that for smaller proteomes a percentage of zero is likelier.
  }
  \label{fig:2}
\end{figure}

@@

We assume that epitopes 
derived from TMHs are actually presented by MHC-I and MHC-II.
Our analysis (at subsection \ref{subsec:elution_studies}) 
finds that 1.4\% (for MHC-I) and 4.0% (for MHC-II) 
of the epitopes presented 
overlap with a TMH.
It is unsure if this a true finding or a false positive.
One source of noise would stem from imperfect predictions
by the topology prediction software used. 
We have counteracted this by using two different topology prediction
software and confirm these have similar results.
Two biological sources of noise
are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
and peptide fusion \cite{delong2016pathogenic},
two processes that both result in peptide fragments
that did not originate from genome directly.
Note however, that both elution studies have a
methodological bias caused by mass spectronomy,
as hydrophobic fragments are detected less 
by mass spectronomy.

@@

In this experiment we predicted epitopes that overlap with 
TMHs from a human, bacterial and viral proteome,
would these proteins be expressed in a human host.
Bacteria, however have different cell membranes and cell walls, 
hence different structural requirements for a TMH.
Both topology prediction tools were trained to recognize
human TMHs, thus we cannot be sure that
the transmembrane regions predicted in bacterial proteins
are actually part of an integral membrane protein.
For the purpose of this study, we assume the 
error in topology predictions to be unbiased way towards topology.
In other words: that a bacterial TMH is incorrectly
predicted to be absent just as often as it is incorrectly
predicted to be present elsewhere.

@@
